---
title: "Series02-Rstudio-RENDU"
output: html_document
date: "2025-10-01"
---

<br>
<br>
<br>
<br>

## 2.1 Make clear what is a statistical test, and when to use it.
Below are some questions to guide your analysis. You can answer them in your notebook, with a short sentence for each question.

- What is a median? A mean?

A median is the value right in the middle between the maximum and the minimum of the dataset.
A mean is the average of all the values of the dataset.  


- What is a variance? A standard deviation?

The variance is the measure of data spread, the average squared deviation from the mean.
The standard deviation is the square root of the variance.


- What is a normal distribution?

A normal distribution is when the values are symetrical around the mean value. Graphicly, it has the shape of a bell.


- What is a statistical test?

A statistical test is a test that allows to decide of data support or reject the hypothesis H0 that was chosen before the test.


- When to use a statistical test?

When we need to determine of a result is statistically significant or if it's due to chance or hazard.


- What is a parametric test? A non-parametric test?

A parametric test is a test where we assume a specific distribution. On the other end, a non-parametric test is where there are no distribution assumptions.


- What are the assumptions of parametric tests?

Normality, homogeneity of variances, independence of observations.


- What are the assumptions of non-parametric tests?

Often require independent samples and ordinal/continuous data.


- What is a p-value?

The p-value is the probability of obtaining data as extreme as observed, under the null hypothesis.
It allows to know if there's a correlation between two datasets.


- What is the risk of error when using a statistical test?

The risk of error in a statistical test means either rejecting a true null hypothesis (Type I error) or failing to reject a 
Reference: *Statistical Methods* (Snedecor & Cochran, 1989)


- What is the difference between a paired and an unpaired test?

A paired test compare measurements on the the same subjects before and after treatments while an unpaired test compare measurements between different groups.


---


<br>
<br>
<br>
<br>
<br>
<br>

# 2.2 Effect of Treatment Over Time

<br>
<br>

```{r setup, include=FALSE}
# Import of all the libraries I'll need in this homework
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(here)
library(tidyr)
library(corrplot)

```



## 2.2.1 The Data
```{r}
# Load data
data_prepost <- read.csv("Datas/PrePost.csv")
print(data_prepost)
```

<table style="border-collapse:collapse; width:100%; max-width:700px; font-family:Arial, sans-serif;">
  <caption style="caption-side:top; text-align:left; font-weight:600; padding:6px 0;">Tableau des variables mesurées</caption>
  <thead>
    <tr>
      <th style="border:1px solid #ccc; padding:8px; text-align:left; background:#f2f2f2;">Variable</th>
      <th style="border:1px solid #ccc; padding:8px; text-align:left; background:#f2f2f2;">Description</th>
      <th style="border:1px solid #ccc; padding:8px; text-align:left; background:#f2f2f2;">Unit</th>
      <th style="border:1px solid #ccc; padding:8px; text-align:left; background:#f2f2f2;">Possible values</th>
      <th style="border:1px solid #ccc; padding:8px; text-align:left; background:#f2f2f2;">Expected change</th>
    </tr>
  </thead>
  <tbody>
    <tr>
      <td style="border:1px solid #ccc; padding:8px;">perf</td>
      <td style="border:1px solid #ccc; padding:8px;">Performance index</td>
      <td style="border:1px solid #ccc; padding:8px;">N/A</td>
      <td style="border:1px solid #ccc; padding:8px;">125–148</td>
      <td style="border:1px solid #ccc; padding:8px;">Increase</td>
    </tr>
    <tr>
      <td style="border:1px solid #ccc; padding:8px;">time</td>
      <td style="border:1px solid #ccc; padding:8px;">Measurement time</td>
      <td style="border:1px solid #ccc; padding:8px;">categorical</td>
      <td style="border:1px solid #ccc; padding:8px;">Before, After</td>
      <td style="border:1px solid #ccc; padding:8px;">N/A</td>
    </tr>
  </tbody>
</table>


<br>
<br>



## 2.2.2 The analysis


- Does the treatment have an effect?

Yes, the treatment appears to have an effect: performance after rehabilitation is higher compared to before.


- What comparison are you going to make?

We compare the same individuals’ performance before vs. after rehabilitation (paired data).


- Which test will you use? Parametric or non-parametric?

The sataset is normally distributed so I will use a paired t-test (parametric).


- Which graph illustrates the effect of the treatment?

A boxplot (Before vs. After) or a line plot connecting paired observations best illustrates the effect.



```{r}
# Load PrePost data
data_prepost <- read.csv("Datas/PrePost.csv")


# Create ID variable for pairing
data_prepost$ID <- rep(1:(nrow(data_prepost)/2), times = 2)


# Ensure Before → After order
data_prepost$time <- factor(data_prepost$time, levels = c("Before", "After"), ordered = TRUE)


# Separate vectors for paired test
before <- data_prepost$perf[data_prepost$time == "Before"]
after <- data_prepost$perf[data_prepost$time == "After"]


# Paired t-test
paired_test <- t.test(before, after, paired = TRUE)
paired_test
```


```{r}
# Boxplot with forced order

ggplot(data_prepost, aes(x = time, y = perf, fill = time)) +
geom_boxplot() +
scale_fill_manual(values = c("Before" = "steelblue", "After" = "seagreen")) +
labs(title = "Performance Before and After Rehabilitation",
x = "Time",
y = "Performance") +
scale_x_discrete(limits = c("Before", "After")) +
theme_minimal()
```


- What sentence will you write in your thesis to explain your result?

The rehabilitation training significantly improved participants’ performance, as shown by a paired t-test (p < 0.05).


---

<br>
<br>
<br>
<br>
<br>
<br>


# 2.3 Testing Some Stereotypes

<br>
<br>

## 2.3.1 The Data
```{r}
data_snore <- read.table("Datas/snore.txt", header = TRUE)
head(data_snore)
```

